---
title: "Nuclear Explosions Data"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(htmlwidgets)
library(leaflet)
library(leaflegend)
library(purrr)
library(viridis)
library(plotly)
library(scales)
library(forcats)
library(ggplot2)
library(ggridges)
library(ggthemes)
library(vistime)
library(ggpubr)
explosions <- read_csv("data/nuclear_explosions.csv")
```

### Interactive plot displaying the number of nuclear bombs by country per year

```{r}

data <- read.csv ("data/nuclear_explosions.csv")

mutated_country <- data %>%
  mutate(
    country = case_when(
      country == "USA" ~ "United States",  
      country == "UK" ~ "United Kingdom",
      country == "USSR" ~ "Soviet Union",
      country == "FRANCE" ~ "France",
      country == "CHINA" ~ "China",
      country == "INDIA" ~ "India",
      country == "PAKIST" ~ "Pakistan",
      TRUE ~ country  # Keep other country names unchanged
    ),
    name = str_to_title(name),  # Convert bomb names to title case
    avg_yield = (yield_upper + yield_lower) / 2,  # Calculate the average yield
    row_id = row_number()  # Add a unique row identifier
  )

mutated_type <- mutated_country %>%
  mutate(type = recode(
    type,
    "ATMOSPH" = "Atmospheric",
    "UG" = "Underground",
    "BALLOON" = "Balloon Drop",
    "AIRDROP" = "Airplane",
    "Airplane Deployed" = "Airplane",
    "ROCKET" = "Rocket",
    "Rocket Deployed" = "Rocket",
    "TOWER" = "Tower",
    "Tower Deployment" = "Tower",
    "BARGE" = "Barge Boat",
    "Barge Boat Deployment" = "Barge Boat",
    "SURFACE" = "Surface",
    "Surface Deployment" = "Surface",
    "UW" = "Underwater",
    "SHAFT" = "Vertical Shaft",
    "TUNNEL/GALLERY" = "Horizontal Tunnel",
    "CRATER" = "Crater",
    "GALLERY" = "Horizontal Tunnel",
    "MINE" = "Underground Mine",
    "SHAFT/GR" = "Vertical Shaft",
    "SHAFT/LG" = "Vertical Shaft",
    "SHIP" = "Ship",
    "SPACE" = "Space",
    "TUNNEL" = "Horizontal Tunnel",
    "WATER SU" = "Water Surface",
    "WATERSURFACE" = "Water Surface",
    "WATERSUR" = "Water Surface"
  ))

number_bombs <- mutated_type %>%
  group_by(country, year) %>%
  summarize(bombs = sum(n())) %>%
  ungroup()

graph_1 <- number_bombs %>%
  ggplot(aes(x = year, y = bombs, fill = country, text = paste0("Country: ", country, "\nBombs: ", bombs, "\nYear: ", year))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  scale_x_continuous(
    breaks = seq(1945, 2000, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, by = 20)
  ) +
  labs(
    title = "Nuclear Bombs By Country Per Year",
    x = "Year",
    y = "Number of Bombs",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

interactive_plot <- ggplotly(graph_1, tooltip = "text")
interactive_plot

```

### Interactive plot displaying the type of nuclear bomb deployments by country

```{r}

library(dplyr)
library(forcats)

type_bombs <- mutated_type %>%
  group_by(country, type) %>%
  summarize(total_bombs = sum(n()), .groups = "drop") %>%
  mutate(type = reorder(type, -total_bombs))

graph_2 <- type_bombs %>%
  ggplot(aes(y = type, x = total_bombs, fill = country, text = paste0("Country: ", country, "\nBombs: ", total_bombs))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Type of Nuclear Bombs By Country (1945-1998)",
    x = "Number of Bombs",
    y = "Type of Deployment",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "right"
  )

interactive_plot <- ggplotly(graph_2, tooltip = "text")
interactive_plot

```

### Interactive plot displaying the top 10 largest yield nuclear bombs, which were all from either the United States or Soviet Union

```{r, fig.width=12, fig.height=4}

filter_unnamed <- mutated_type %>%
  filter(is.na(name)) %>%  # Filter rows where name is NA
  mutate(
    is_unnamed = TRUE,  # Flag for unnamed bombs
    unnamed_id = row_number(),  # Sequential numbering for unnamed bombs
    name = paste0("Unnamed Bomb ", unnamed_id)  # Assign sequential names
  )

mutated_type <- mutated_type %>%
  left_join(
    filter_unnamed %>% select(row_id, name),  # Select only row_id and the new names
    by = "row_id"
  ) %>%
  mutate(name = coalesce(name.y, name.x)) %>%  # Replace NA names with new sequential names
  select(-name.x, -name.y)  # Clean up temporary columns

top_10_yield <- mutated_type %>%
  arrange(desc(avg_yield)) %>%  # Order by highest yield
  slice_head(n = 10)  # Select the top 10 bombs

graph_3 <- top_10_yield %>%
  ggplot(aes(
    x = reorder(name, avg_yield),  # Order names by yield
    y = avg_yield,
    fill = country,
    text = paste0(
      "Year: ", year,
      "\nAverage Yield: ", round(avg_yield, 1), " kilotons"
    )
  )) +
  geom_bar(stat = "identity") +
    scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  coord_flip() +  # Flip axes for readability
  labs(
    title = "Top 10 Largest Nuclear Bombs",
    x = "Bomb Name",
    y = "Average Yield (kilotons)",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
  )

interactive_plot <- ggplotly(graph_3, tooltip = "text")
interactive_plot

```

### Interactive plot with the largest nuclear bombs for each country

```{r, fig.width=12, fig.height=2}

# Find the largest bomb for each country
largest_bomb_per_country <- mutated_type %>%
  group_by(country) %>%
  filter(avg_yield == max(avg_yield, na.rm = TRUE)) %>%  # Select the bomb with the highest yield
  ungroup()  # Remove grouping for further operations

# Plot the largest bombs
graph_4 <- largest_bomb_per_country %>%
  ggplot(aes(
    x = reorder(country, avg_yield),  # Order countries by yield
    y = avg_yield,
    fill = country,
    text = paste0(
      "Name: ", name,
      "\nAverage Yield: ", round(avg_yield, 1), " kilotons",
      "\nYear: ", year
    )
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Largest Nuclear Bomb From Each Country",
    x = "Country",
    y = "Average Yield (kilotons)",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "none"  # Hide the legend since the bars already indicate countries
  )

interactive_plot <- ggplotly(graph_4, tooltip = "text")
interactive_plot

```

### Interactive plot displaying the top 10 most common regions where the bombs were deployed

```{r, fig.width=12, fig.height=4}

# Mutate regions with descriptive names and country locations
mutated_regions <- mutated_type %>%
  mutate(
    region_desc = case_when(
      region == "ALAMOGORDO" ~ "Alamogordo, New Mexico, USA",
      region == "AMCHITKA AK" ~ "Amchitka Island, Alaska, USA",
      region == "ARKHAN RUSS" ~ "Arkhangelsk, Russia",
      region == "ASTRAK RUSS" ~ "Astrakhan, Russia",
      region == "AZGIE KAZAKH" ~ "Azgir, Kazakhstan",
      region == "AZGIR" ~ "Azgir, Kazakhstan",
      region == "AZGIR KAZAKH" ~ "Azgir, Kazakhstan",
      region == "BASHKI RUSS" ~ "Bashkiria, Russia",
      region == "BASHKIR RUSS" ~ "Bashkiria, Russia",
      region == "BIKINI" ~ "Bikini Atoll, Marshall Islands, USA",
      region == "C. NEVADA" ~ "Central Nevada, USA",
      region == "CARLSBAD NM" ~ "Carlsbad, New Mexico, USA",
      region == "CHAGAI" ~ "Chagai Hills, Pakistan",
      region == "CHITA RUSS" ~ "Chita, Russia",
      region == "CHRISTMAS IS" ~ "Christmas Island (Kiritimati), British Territory",
      region == "EMU AUSTR" ~ "Emu Field, Australia",
      region == "ENEWETAK" ~ "Enewetak Atoll, Marshall Islands, USA",
      region == "FALLON NV" ~ "Fallon, Nevada, USA",
      region == "FANGATAUFA" ~ "Fangataufa Atoll, French Polynesia, France",
      region == "FANGATAUFAA" ~ "Fangataufa Atoll, French Polynesia, France",
      region == "FARMINGT NM" ~ "Farmington, New Mexico, USA",
      region == "GRAND V CO" ~ "Grand Valley, Colorado, USA",
      region == "HATTIESB MS" ~ "Hattiesburg, Mississippi, USA",
      region == "HATTIESE MS" ~ "Hattiesburg, Mississippi, USA",
      region == "HIROSHIMA" ~ "Hiroshima, Japan",
      region == "HTR RUSS" ~ "Hot Range, Russia",
      region == "HURUROA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "IN ECKER ALG" ~ "In Ekker, Algeria (French Test)",
      region == "IRKUTS RUSS" ~ "Irkutsk, Russia",
      region == "JAKUTS RUSE" ~ "Yakutsk, Russia",
      region == "JAKUTS RUSS" ~ "Yakutsk, Russia",
      region == "JOHNSTON IS" ~ "Johnston Atoll, Pacific Ocean, USA",
      region == "KALMYK RUSS" ~ "Kalmykia, Russia",
      region == "KAZAKH" ~ "Kazakhstan (Semipalatinsk Test Site), USSR",
      region == "KAZAKHSTAN" ~ "Kazakhstan (Semipalatinsk Test Site), USSR",
      region == "KEMERO RUSS" ~ "Kemerovo, Russia",
      region == "KHARAN" ~ "Kharan, Pakistan",
      region == "KOMI RUSS" ~ "Komi Republic, Russia",
      region == "KRASNO RUSS" ~ "Krasnoyarsk, Russia",
      region == "KZ RUSS" ~ "Kazakhstan, USSR",
      region == "LOP NOR" ~ "Lop Nur, Xinjiang, China",
      region == "MALDEN IS" ~ "Malden Island, British Territory",
      region == "MANGY KAZAKH" ~ "Mangyshlak Peninsula, Kazakhstan, USSR",
      region == "MARALI AUSTR" ~ "Maralinga, Australia",
      region == "MARY TURKMEN" ~ "Mary, Turkmenistan, USSR",
      region == "MELLIS NV" ~ "Nellis Air Force Base, Nevada, USA",
      region == "MONTEB AUSTR" ~ "Monte Bello Islands, Australia",
      region == "MTR RUSS" ~ "Moscow Test Range, Russia",
      region == "MUEUEOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURM RUSS" ~ "Murmansk, Russia",
      region == "MURUEOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURUHOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURUROA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "N2 RUSS" ~ "Unknown, Russia",
      region == "NAGASAKI" ~ "Nagasaki, Japan",
      region == "NELLIS NV" ~ "Nellis Air Force Base, Nevada, USA",
      region == "NTS" ~ "Nevada Test Site, USA",
      region == "NZ RUSS" ~ "Novaya Zemlya, Russia",
      region == "OFFUSWCOAST" ~ "Off the Southwest Coast, USA",
      region == "ORENBG RUSS" ~ "Orenburg, Russia",
      region == "PACIFIC" ~ "Pacific Ocean, USA",
      region == "PAMUK UZBEK" ~ "Pamuk, Uzbekistan, USSR",
      region == "PERM RUSS" ~ "Perm, Russia",
      region == "POKHRAN" ~ "Pokhran, Rajasthan, India",
      region == "REGGANE ALG" ~ "Reggane, Algeria (French Test)",
      region == "RIFLE CO" ~ "Rifle, Colorado, USA",
      region == "S. ATLANTIC" ~ "South Atlantic Ocean, USA",
      region == "S.ATLANTIC" ~ "South Atlantic Ocean, USA",
      region == "SEMI KAZAKH" ~ "Semipalatinsk, Kazakhstan (USSR)",
      region == "STAVRO RUSS" ~ "Stavropol, Russia",
      region == "TUYMEN RUSS" ~ "Tyumen, Russia",
      region == "TYUMEN RUSS" ~ "Tyumen, Russia",
      region == "UKEAINE" ~ "Ukraine (USSR)",
      region == "UKRAINE" ~ "Ukraine (USSR)",
      region == "URAL RUSS" ~ "Ural Mountains, Russia",
      region == "UZBEK" ~ "Uzbekistan, USSR",
      region == "W KAZAKH" ~ "Western Kazakhstan, USSR",
      region == "W MURUROA" ~ "Western Mururoa Atoll, French Polynesia",
      region == "WSW MURUROA" ~ "West-Southwest Mururoa Atoll, French Polynesia",
      TRUE ~ "Unknown/Unmapped Region"  # Default for regions not explicitly listed
    )
  )

# Count the number of bombs per region
top_regions <- mutated_regions %>%
  group_by(region_desc, country) %>%
  summarize(total_bombs = n(), .groups = "drop") %>%  # Count bombs in each region
  arrange(desc(total_bombs)) %>%  # Sort by the number of bombs
  slice_head(n = 10)  # Select the top 10 regions

# Create the bar plot
graph_5 <- top_regions %>%
  ggplot(aes(
    x = reorder(region_desc, total_bombs),  # Order regions by bomb count
    y = total_bombs,
    fill = country,
    text = paste0(
      "Region: ", region_desc,
      "\nCountry: ", country,
      "\nNumber of Bombs: ", total_bombs
    )
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Top 10 Most Common Regions for Bomb Drops",
    x = "Region",
    y = "Number of Bombs",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16)
  )

interactive_plot <- ggplotly(graph_5, tooltip = "text")
interactive_plot

```

### World map of nuclear explosions

```{r}
# Owen's graph

# Github link: https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-20

avg <- function(x, y) {
  num <- (x + y) / 2
}

# Data tidying
explosions2 <- explosions %>% 
  # Rename purpose to something more human readable
  mutate(map_purpose = case_when(
    purpose == "COMBAT" ~ "WWII bombs dropped over Japan",
    purpose == "FMS" ~ "Soviet test studying phenomenon of nuclear explosion",
    purpose == "ME" ~ "Military Exercise",
    purpose == "PNE" ~ "Peaceful nuclear explosion",
    purpose == "SAM" ~ "Soviet accident test",
    purpose == "TRANSP" ~ "Transportation-storage purpose",
    purpose == "WE" ~ "Evaluate effects of nuclear detonation on various targets",
    purpose == "WR" ~ "Weapons development program",
    TRUE ~ purpose)) %>% 
  mutate(map_type = case_when(
    type == "ATMOSPH" ~ "Atmospheric",
    type == "UG" ~ "Underground",
    type == "BALLOON" ~ "Balloon drop",
    type == "AIRDROP" ~ "Airplane drop",
    type == "ROCKET" ~ "Rocket",
    type == "TOWER" ~ "Constructed tower",
    type == "WATERSURFACE" ~ "Surface of a water body",
    type == "BARGE" ~ "Barge boat",
    type == "SURFACE" ~ "Surface or shallow crater",
    type == "UW" ~ "Underwater",
    type == "SHAFT" ~ "Vertical shaft underground",
    type == "TUNNEL/GALLERY" ~ "Horizontal tunnel",
    TRUE ~ type)) %>% 
  mutate(average = map2_dbl(.x = yield_lower, .y = yield_upper, 
                        ~ avg(.x, .y))) %>% 
  mutate(scaled_average = average / 1000) %>%
  mutate(plot_scaled_average = ifelse(scaled_average < 1, 1, scaled_average)) %>% 
  mutate(popup_info = paste0("<b>Estimated yield (Mt TNT): </b>", scaled_average, "<br/>",
                             "<b>Deployment Method: </b>", map_type, "<br/>",
                             "<b>Purpose: </b>", map_purpose, "<br/>",
                             "<b>Country: </b>", country, "<br/>",
                             "<b>Region: </b>", region, "<br/>",
                             "<b>Year: </b>", year, "<br/>"))

# Create a palette for explosion sizes
pal <- colorNumeric(palette = 'inferno', 
                   domain = explosions2$plot_scaled_average,
                   na.color = NA)

leaflet() %>% 
  addTiles() %>% 
  addProviderTiles("Thunderforest.Landscape") %>% 
  addCircleMarkers(data = explosions2, lat = ~ latitude, lng = ~ longitude,
                   radius = ~ plot_scaled_average, 
                   popup = ~ popup_info,
                   stroke = FALSE, fillOpacity = 0.5,
                   color = ~ pal(plot_scaled_average)) %>% 
  addLegend(data = explosions2,
            pal = pal,
            values = ~ plot_scaled_average,
            title = "Worldwide nuclear explosions <br/> Megatonnes of TNT")
```
***
This map shows each nuclear explosion since 1945. The sizes of the circles are scaled and colored based on the estimated yield in megatonnes of TNT. Each circle has a popup that shows info on the estimated yield, purpose of explosion, country, region, and year.


### Explosions Per Period

```{r}

major_events <- read.csv("data/Major Events.csv")

#Generate Separate Datasets by Periods

explosion_event <- major_events %>%
  full_join(explosions, by = "year")
explosions_1945_1954 <- explosion_event %>%
  filter(year <= 1954) %>%
  filter(!is.na(region)) 

region_1945_1954 <- explosions_1945_1954 %>%
  group_by(region) %>%
  summarize(count = n())%>%
  right_join(explosions_1945_1954, by = 'region')

explosions_1953_1959 <- explosion_event %>%
  filter(year >= 1953 & year <= 1959) %>%
  filter(!is.na(region))

region_1953_1959 <- explosions_1953_1959 %>%
  group_by(region) %>%
  summarize(count = n()) %>%
  right_join(explosions_1953_1959, by = 'region')
explosions_1955_1975 <- explosion_event %>%
  filter(year >= 1955 & year <= 1975) %>%
  filter(!is.na(region))

region_1955_1975 <- explosions_1955_1975 %>%
  group_by(region) %>%
  summarize(count = n())%>%
  right_join(explosions_1955_1975, by = 'region') 
explosions_1979_1989 <- explosion_event %>%
  filter(year >= 1979 & year <= 1989) %>%
  filter(!is.na(region))

region_1979_1989 <- explosions_1979_1989 %>%
  group_by(region) %>%
  summarize(count = n())%>%
  right_join(explosions_1979_1989, by = 'region')

pal1<- leaflet::colorNumeric("viridis", domain = region_1945_1954$count)

pal2<- leaflet::colorNumeric("viridis", domain = region_1953_1959$count)

pal3<- leaflet::colorNumeric("viridis", domain = region_1955_1975$count)

pal4<- leaflet::colorNumeric("viridis", domain = region_1979_1989$count)


#graph with leaflet to generate layers by period
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = region_1945_1954, ~longitude, ~latitude , radius = 3, color = ~pal1(count), fillOpacity = 0.7,group = "1945-1954")%>%
  addLegend(data = region_1945_1954,
            position = "bottomright",
            pal = pal1, values = ~count,
            title = "# of Events",
            opacity = 1, group="1945-1954") %>%
  addCircleMarkers(data = region_1953_1959, ~longitude, ~latitude , radius = 3, color = ~pal2(count), fillOpacity = 0.7,group = "1953-1959")%>%
  addLegend(data = region_1953_1959,
            position = "bottomright",
            pal = pal2, values = ~count,
            title = "# of Events",
            opacity = 1, group="1953-1959") %>%
  addCircleMarkers(data = region_1955_1975, ~longitude, ~latitude , radius = 3, color = ~pal3(count), fillOpacity = 0.7,group = "1955-1975")%>%
  addLegend(data = region_1955_1975,
            position = "bottomright",
            pal = pal3, values = ~count,
            title = "# of Events",
            opacity = 1, group = "1955-1975") %>%
  addCircleMarkers(data = region_1979_1989, ~longitude, ~latitude , radius = 3, color = ~pal4(count), fillOpacity = 0.7,group = "1979-1989")%>%
  addLegend(data = region_1979_1989,
            position = "bottomright",
            pal = pal4, values = ~count,
            title = "# of Events",
            opacity = 1, group = "1979-1989") %>%
  addLayersControl(overlayGroups = c("1945-1954", "1953-1959", "1955-1975", "1979-1989"), position = "topright", options = layersControlOptions(collapsed = F)) %>%
  hideGroup(c("1953-1959", "1955-1975", "1979-1989"))



```

***
The following map shows the number of nuclear tests across the globe during four periods:

-The End of WWII through the end of the Korean War (1945-1954)

-The Cuban Revolution (1953-1959)

-The Vietnam War (1955-1975)

-The Iran-Iraq War and Soviet-Afghan War (1979-1989)



### Nuclear Explosions per Country

```{r, echo=FALSE, fig.width=6.5, fig.height=7.5}

#plot looking at trend of each country from 1945 to 1998

counts_by_year <- explosion_event %>%
  filter(!is.na(region)) %>%
  group_by(year, country) %>%
  summarize(n = n()) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(country = recode(country, CHINA = "China", FRANCE = "France", INDIA = "India", PAKIST = "Pakistan"))

PAK_events <- counts_by_year %>%
  filter(country == "Pakistan")

nukes_by_year <- ggplot(data=counts_by_year, aes (x = year, y = n, color = country)) +
  geom_line(linewidth = 1) +
  geom_point(data = PAK_events, color="orange") +
  theme_bw() +
  labs(y = "Number of Nuclear Explosion Events",x = "Year") +
  scale_x_continuous(breaks = seq(1945, 2000, 5), limits = c(1945, 1998)) +
  scale_color_manual(breaks = c("USA", "USSR", "UK", "China", "France", "India", "Pakistan"), values = c("black", "red", "steelblue", "hotpink3", "blue", "green3", "orange")) +
  theme(panel.background = element_blank(),,panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),plot.background = element_blank()) +
  theme(legend.position = "top", legend.title = element_blank()) +
  guides(colour = guide_legend(nrow = 1)) 

timeline <- data.frame(Position = rep(c("", "", "", "", "", "", "", "", ""), each = 1),
                   Name = c("World War II", "Chinese Civil War", "First IndoChina War", "Korean War", "Cuban Revolution", "Vietnam War", "Lebanese Civil War", "Soviet-Afghan War", "Iran-Iraq War"),
                   start = c("1945-01-01", "1945-01-01", "1946-01-01", "1951-01-01", "1953-01-01", "1955-01-01", "1975-01-01", "1979-01-01", "1980-01-01" ),
                   end = c("1945-12-31", "1949-12-31", "1954-12-31", "1954-12-31", "1959-12-31", "1975-12-31", "1990-12-31", "1989-12-31", "1988-12-31"),
                   color = c("lightgrey", "azure2", "skyblue", "skyblue2", "steelblue","steelblue4",  "blue",  "blue3", "navy"))

wars<- gg_vistime(timeline, col.event = "Position", col.group = "Name") +
  scale_x_datetime(date_breaks = "5 years", date_labels = "%Y", limits = as.POSIXct(c("1945-01-01", "2000-01-01"))) +
  theme_tufte() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  theme()

wars <- ggarrange(nukes_by_year, wars, nrow = 2, heights = c(3, 1.5), align = 'v')

wars

```

<div>
The following plot shows conflicts
<div>
from 1945-1998 and the amount
<div> 
of nuclear bombs tested by the 
<div>
different countries during 
<div>
The Cold War and the many conflicts 
<div>
that occurred during that time.
<div>



### Nuclear Explosions by Purpose

```{r}
purpose <- explosions %>%
  group_by(purpose) %>%
  summarise(number_instances = n(), .groups = "drop")
```
```{r}
top_5_purposes <- purpose %>%
  arrange(desc(number_instances)) %>%
  slice_head(n = 5)
```
```{r}
top_5_purposes_data <- explosions %>%
  dplyr::filter(purpose %in% top_5_purposes$purpose) %>%
  # Join with the `top_5_purposes` to get `number_instances` from the purpose dataset
  left_join(top_5_purposes, by = "purpose") %>%
  select(purpose, number_instances, country, yield_upper)
```
```{r}
top_5_purposes_data <-  top_5_purposes_data %>%
  mutate(
    purpose = recode(purpose,
                     "FMS" = "Soviet test",
                     "PNE" = "Peaceful nuclear explosion",
                     "SE" = "Accidental safety testing",
                     "WE" = "Target nuclear detonation",
                     "WR" = "Weapons development")
  )
```

```{r,fig.width=10, fig.height=5}
ggplot(top_5_purposes_data, aes(x = number_instances, y = purpose, fill = purpose, alpha = 0.5)) +
  geom_density_ridges() +
  scale_x_log10() +  # Apply log10 transformation to the x-axis
  scale_fill_manual(values = c("yellow", "green", "purple", "orange", "red")) +  # Custom colors
  scale_x_continuous(labels = label_number()) +
  labs(title = 'Purpose of Nuclear Testing versus Number of Times', x = 'Number of Times', y = 'Purpose of Nuclear Testing') +
  theme_ridges() +
  theme(legend.position = "none",
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5, angle = 90))
```

***

This plot illustrates the distribution of nuclear tests by purpose using a density ridge plot. The x-axis represents the number of times a particular nuclear testing purpose occurred and the y-axis shows the top five most common purposes for testing.












### Arielle's Graph
```{r}
# Add code here


```
***
Add description here
