---
title: "Final_Project"
author: "Alex Alon, Owen Bevis, Arielle Glass, Susan Gogolski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      error = FALSE)
```

```{r, echo=TRUE}
```

```



### Load libraries and data
```{r libraries}
library(tidyverse)
library(htmlwidgets)
library(leaflet)
library(leaflegend)
library(purrr)
library(viridis)
```

```{r data}
explosions <- read_csv("data/nuclear_explosions.csv")
```

```{r AL}
# Alex's graph


```

```{r OB}
# Owen's graph

# Github link: https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-20

avg <- function(x, y) {
  num <- (x + y) / 2
}

# Data tidying
explosions2 <- explosions %>% 
  # Rename 'purpose' as indicated in data README
  mutate(map_purpose = if_else(purpose == "COMBAT",
                               "WWII bombs dropped over Japan", purpose),
         map_purpose = if_else(purpose == "FMS",
                               "Soviet test studying phenomenon of nuclear explosion", purpose),
         map_purpose = if_else(purpose == "ME",
                               "Military Exercise", purpose),
         map_purpose = if_else(purpose == "PNE",
                               "Peaceful nuclear explosion", purpose),
         map_purpose = if_else(purpose == "SAM",
                               "Soviet accident test", purpose),
         map_purpose = if_else(purpose == "TRANSP",
                               "Transportation-storage purpose", purpose),
         map_purpose = if_else(purpose == "WE", 
                               "Evaluate effects of nuclear detonation on various targets", purpose),
         map_purpose = if_else(purpose == "WR",
                               "Weapons development program", purpose)) %>% 
  mutate(average = map2_dbl(.x = yield_lower, .y = yield_upper, 
                        ~ avg(.x, .y))) %>% 
  mutate(scaled_average = average / 1000) %>%
  mutate(plot_scaled_average = ifelse(scaled_average < 1, 1, scaled_average)) %>% 
  mutate(popup_info = paste0("<b>Estimated yield (Mt TNT): </b>", scaled_average, "<br/>",
                             "<b>Purpose: </b>", map_purpose, "<br/>",
                             "<b>Country: </b>", country, "<br/>",
                             "<b>Region: </b>", region, "<br/>",
                             "<b>Year: </b>", year, "<br/>"))

# Create a palette for explosion sizes
pal <- colorNumeric(palette = 'viridis', 
                   domain = explosions2$plot_scaled_average,
                   na.color = NA)
explosions2
# To do:
# Fix popups
# Fix title
# Fix scale so that values < 1 will still show up as 1

leaflet() %>% 
  addTiles() %>% 
  addProviderTiles("Thunderforest.Landscape") %>% 
  addCircleMarkers(data = explosions2, lat = ~ latitude, lng = ~ longitude,
                   radius = ~ plot_scaled_average, 
                   popup = ~ popup_info,
                   stroke = FALSE, fillOpacity = 0.5,
                   color = ~ pal(plot_scaled_average)) %>% 
  addLegend(data = explosions2,
            pal = pal,
            values = ~ plot_scaled_average,
            title = "Worldwide nuclear explosions <br/> Megatonnes of TNT")
```


```{r AG}
# Arielle's graph


```

```{r SG}
# Susan's graph
```{r}
library(ggplot2)
library(tidyverse)
library(conflicted)
library(ggridges)
library(plotly)
library(readr)
library(dplyr)
library(RColorBrewer) 
library(forcats)
library(scales)
library(stringr)
```
```{r}
explosions <- read_csv('/Users/s/Desktop/final_project/data/nuclear_explosions.csv')

```
# Calculate total number of instances for each purpose
```{r}
purpose <- explosions %>%
  group_by(purpose) %>%
  summarise(number_instances = n(), .groups = "drop")
```
# Select the top 5 purposes bases on the number of instancee
```{r}
top_5_purposes <- purpose %>%
  arrange(desc(number_instances)) %>%
  slice_head(n = 5)
```
# Join with the orginal dataset to get corresponding `country` and `yield_upper`
# Here, we use `left_join` to keep all columns from the original `purpose` data

```{r}
# Join the top 5 purposes with the explosions data
top_5_purposes_data <- explosions %>%
  dplyr::filter(purpose %in% top_5_purposes$purpose) %>%
  # Join with the `top_5_purposes` to get `number_instances` from the purpose dataset
  left_join(top_5_purposes, by = "purpose") %>%
  select(purpose, number_instances, country, yield_upper, year)
```
```{r}
# Replace abbreviations with full words
top_5_purposes_data <-  top_5_purposes_data %>%
  mutate(
    purpose = recode(purpose,
                     "FMS" = "Soviet test",
                     "PNE" = "Peaceful nuclear explosion",
                     "SE" = "Accidental safety testing",
                     "WE" = "Target nuclear detonation",
                     "WR" = "Weapons development")
  )
```

```{r,fig.width=10, fig.height=5}
ggplot(top_5_purposes_data, aes(x = number_instances, y = purpose, fill = purpose, alpha = 0.5)) +
  geom_density_ridges() +
  scale_x_log10() +  # Apply log10 transformation to the x-axis
  scale_fill_manual(values = c("yellow", "green", "purple", "orange", "red")) +  # Custom colors
  scale_x_continuous(labels = label_number()) +
  labs(title = 'Purpose of Nuclear Testing versus Number of Times', x = 'Number of Times', y = 'Purpose of Nuclear Testing') +
  theme_ridges() +
  theme(legend.position = "none",
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5, angle = 90))
```



