---
title: "Final project- AG"
author: "Arielle Glass"
date: "2024-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(plotly)
library(ggpubr)
```

Interactive plot displaying the number of nuclear bombs by country per year

```{r}

data <- read.csv ("./data/nuclear_explosions.csv")

mutated_country <- data %>%
  mutate(
    country = case_when(
      country == "USA" ~ "United States",  
      country == "UK" ~ "United Kingdom",
      country == "USSR" ~ "Soviet Union",
      country == "FRANCE" ~ "France",
      country == "CHINA" ~ "China",
      country == "INDIA" ~ "India",
      country == "PAKIST" ~ "Pakistan",
      TRUE ~ country  # Keep other country names unchanged
    ),
    name = str_to_title(name),  # Convert bomb names to title case
    avg_yield = (yield_upper + yield_lower) / 2,  # Calculate the average yield
    row_id = row_number()  # Add a unique row identifier
  )

mutated_type <- mutated_country %>%
  mutate(type = recode(
    type,
    "ATMOSPH" = "Atmospheric",
    "UG" = "Underground",
    "BALLOON" = "Balloon Drop",
    "AIRDROP" = "Airplane",
    "Airplane Deployed" = "Airplane",
    "ROCKET" = "Rocket",
    "Rocket Deployed" = "Rocket",
    "TOWER" = "Tower",
    "Tower Deployment" = "Tower",
    "BARGE" = "Barge Boat",
    "Barge Boat Deployment" = "Barge Boat",
    "SURFACE" = "Surface",
    "Surface Deployment" = "Surface",
    "UW" = "Underwater",
    "SHAFT" = "Vertical Shaft",
    "TUNNEL/GALLERY" = "Horizontal Tunnel",
    "CRATER" = "Crater",
    "GALLERY" = "Horizontal Tunnel",
    "MINE" = "Underground Mine",
    "SHAFT/GR" = "Vertical Shaft",
    "SHAFT/LG" = "Vertical Shaft",
    "SHIP" = "Ship",
    "SPACE" = "Space",
    "TUNNEL" = "Horizontal Tunnel",
    "WATER SU" = "Water Surface",
    "WATERSURFACE" = "Water Surface",
    "WATERSUR" = "Water Surface"
  ))

number_bombs <- mutated_type %>%
  group_by(country, year) %>%
  summarize(bombs = sum(n())) %>%
  ungroup()

graph_1 <- number_bombs %>%
  ggplot(aes(x = year, y = bombs, fill = country, text = paste0("Country: ", country, "\nBombs: ", bombs, "\nYear: ", year))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  scale_x_continuous(
    breaks = seq(1945, 2000, by = 5)
  ) +
  scale_y_continuous(
    breaks = seq(0, 200, by = 20)
  ) +
  labs(
    title = "Nuclear Bombs By Country Per Year",
    x = "Year",
    y = "Number of Bombs",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

interactive_plot <- ggplotly(graph_1, tooltip = "text")
interactive_plot

```
Interactive plot displaying the type of nuclear bomb deployments by country

```{r}

library(dplyr)
library(forcats)

type_bombs <- mutated_type %>%
  group_by(country, type) %>%
  summarize(total_bombs = sum(n()), .groups = "drop") %>%
  mutate(type = reorder(type, -total_bombs))

graph_2 <- type_bombs %>%
  ggplot(aes(y = type, x = total_bombs, fill = country, text = paste0("Country: ", country, "\nBombs: ", total_bombs))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Type of Nuclear Bombs By Country (1945-1998)",
    x = "Number of Bombs",
    y = "Type of Deployment",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "right"
  )

interactive_plot <- ggplotly(graph_2, tooltip = "text")
interactive_plot

```
Interactive plot displaying the top 10 largest yield nuclear bombs, which were all from either the United States or Soviet Union

```{r, fig.width=12, fig.height=4}

filter_unnamed <- mutated_type %>%
  filter(is.na(name)) %>%  # Filter rows where name is NA
  mutate(
    is_unnamed = TRUE,  # Flag for unnamed bombs
    unnamed_id = row_number(),  # Sequential numbering for unnamed bombs
    name = paste0("Unnamed Bomb ", unnamed_id)  # Assign sequential names
  )

mutated_type <- mutated_type %>%
  left_join(
    filter_unnamed %>% select(row_id, name),  # Select only row_id and the new names
    by = "row_id"
  ) %>%
  mutate(name = coalesce(name.y, name.x)) %>%  # Replace NA names with new sequential names
  select(-name.x, -name.y)  # Clean up temporary columns

top_10_yield <- mutated_type %>%
  arrange(desc(avg_yield)) %>%  # Order by highest yield
  slice_head(n = 10)  # Select the top 10 bombs

graph_3 <- top_10_yield %>%
  ggplot(aes(
    x = reorder(name, avg_yield),  # Order names by yield
    y = avg_yield,
    fill = country,
    text = paste0(
      "Year: ", year,
      "\nAverage Yield: ", round(avg_yield, 1), " kilotons"
    )
  )) +
  geom_bar(stat = "identity") +
    scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  coord_flip() +  # Flip axes for readability
  labs(
    title = "Top 10 Largest Nuclear Bombs",
    x = "Bomb Name",
    y = "Average Yield (kilotons)",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
  )

interactive_plot <- ggplotly(graph_3, tooltip = "text")
interactive_plot

```
Interactive plot with the largest nuclear bombs for each country

```{r, fig.width=12, fig.height=2}

# Find the largest bomb for each country
largest_bomb_per_country <- mutated_type %>%
  group_by(country) %>%
  filter(avg_yield == max(avg_yield, na.rm = TRUE)) %>%  # Select the bomb with the highest yield
  ungroup()  # Remove grouping for further operations

# Plot the largest bombs
graph_4 <- largest_bomb_per_country %>%
  ggplot(aes(
    x = reorder(country, avg_yield),  # Order countries by yield
    y = avg_yield,
    fill = country,
    text = paste0(
      "Name: ", name,
      "\nAverage Yield: ", round(avg_yield, 1), " kilotons",
      "\nYear: ", year
    )
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Largest Nuclear Bomb From Each Country",
    x = "Country",
    y = "Average Yield (kilotons)",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "none"  # Hide the legend since the bars already indicate countries
  )

interactive_plot <- ggplotly(graph_4, tooltip = "text")
interactive_plot

```
Interactive plot displaying the top 10 most common regions where the bombs were deployed

```{r, fig.width=12, fig.height=4}

# Mutate regions with descriptive names and country locations
mutated_regions <- mutated_type %>%
  mutate(
    region_desc = case_when(
      region == "ALAMOGORDO" ~ "Alamogordo, New Mexico, USA",
      region == "AMCHITKA AK" ~ "Amchitka Island, Alaska, USA",
      region == "ARKHAN RUSS" ~ "Arkhangelsk, Russia",
      region == "ASTRAK RUSS" ~ "Astrakhan, Russia",
      region == "AZGIE KAZAKH" ~ "Azgir, Kazakhstan",
      region == "AZGIR" ~ "Azgir, Kazakhstan",
      region == "AZGIR KAZAKH" ~ "Azgir, Kazakhstan",
      region == "BASHKI RUSS" ~ "Bashkiria, Russia",
      region == "BASHKIR RUSS" ~ "Bashkiria, Russia",
      region == "BIKINI" ~ "Bikini Atoll, Marshall Islands, USA",
      region == "C. NEVADA" ~ "Central Nevada, USA",
      region == "CARLSBAD NM" ~ "Carlsbad, New Mexico, USA",
      region == "CHAGAI" ~ "Chagai Hills, Pakistan",
      region == "CHITA RUSS" ~ "Chita, Russia",
      region == "CHRISTMAS IS" ~ "Christmas Island (Kiritimati), British Territory",
      region == "EMU AUSTR" ~ "Emu Field, Australia",
      region == "ENEWETAK" ~ "Enewetak Atoll, Marshall Islands, USA",
      region == "FALLON NV" ~ "Fallon, Nevada, USA",
      region == "FANGATAUFA" ~ "Fangataufa Atoll, French Polynesia, France",
      region == "FANGATAUFAA" ~ "Fangataufa Atoll, French Polynesia, France",
      region == "FARMINGT NM" ~ "Farmington, New Mexico, USA",
      region == "GRAND V CO" ~ "Grand Valley, Colorado, USA",
      region == "HATTIESB MS" ~ "Hattiesburg, Mississippi, USA",
      region == "HATTIESE MS" ~ "Hattiesburg, Mississippi, USA",
      region == "HIROSHIMA" ~ "Hiroshima, Japan",
      region == "HTR RUSS" ~ "Hot Range, Russia",
      region == "HURUROA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "IN ECKER ALG" ~ "In Ekker, Algeria (French Test)",
      region == "IRKUTS RUSS" ~ "Irkutsk, Russia",
      region == "JAKUTS RUSE" ~ "Yakutsk, Russia",
      region == "JAKUTS RUSS" ~ "Yakutsk, Russia",
      region == "JOHNSTON IS" ~ "Johnston Atoll, Pacific Ocean, USA",
      region == "KALMYK RUSS" ~ "Kalmykia, Russia",
      region == "KAZAKH" ~ "Kazakhstan (Semipalatinsk Test Site), USSR",
      region == "KAZAKHSTAN" ~ "Kazakhstan (Semipalatinsk Test Site), USSR",
      region == "KEMERO RUSS" ~ "Kemerovo, Russia",
      region == "KHARAN" ~ "Kharan, Pakistan",
      region == "KOMI RUSS" ~ "Komi Republic, Russia",
      region == "KRASNO RUSS" ~ "Krasnoyarsk, Russia",
      region == "KZ RUSS" ~ "Kazakhstan, USSR",
      region == "LOP NOR" ~ "Lop Nur, Xinjiang, China",
      region == "MALDEN IS" ~ "Malden Island, British Territory",
      region == "MANGY KAZAKH" ~ "Mangyshlak Peninsula, Kazakhstan, USSR",
      region == "MARALI AUSTR" ~ "Maralinga, Australia",
      region == "MARY TURKMEN" ~ "Mary, Turkmenistan, USSR",
      region == "MELLIS NV" ~ "Nellis Air Force Base, Nevada, USA",
      region == "MONTEB AUSTR" ~ "Monte Bello Islands, Australia",
      region == "MTR RUSS" ~ "Moscow Test Range, Russia",
      region == "MUEUEOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURM RUSS" ~ "Murmansk, Russia",
      region == "MURUEOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURUHOA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "MURUROA" ~ "Mururoa Atoll, French Polynesia, France",
      region == "N2 RUSS" ~ "Unknown, Russia",
      region == "NAGASAKI" ~ "Nagasaki, Japan",
      region == "NELLIS NV" ~ "Nellis Air Force Base, Nevada, USA",
      region == "NTS" ~ "Nevada Test Site, USA",
      region == "NZ RUSS" ~ "Novaya Zemlya, Russia",
      region == "OFFUSWCOAST" ~ "Off the Southwest Coast, USA",
      region == "ORENBG RUSS" ~ "Orenburg, Russia",
      region == "PACIFIC" ~ "Pacific Ocean, USA",
      region == "PAMUK UZBEK" ~ "Pamuk, Uzbekistan, USSR",
      region == "PERM RUSS" ~ "Perm, Russia",
      region == "POKHRAN" ~ "Pokhran, Rajasthan, India",
      region == "REGGANE ALG" ~ "Reggane, Algeria (French Test)",
      region == "RIFLE CO" ~ "Rifle, Colorado, USA",
      region == "S. ATLANTIC" ~ "South Atlantic Ocean, USA",
      region == "S.ATLANTIC" ~ "South Atlantic Ocean, USA",
      region == "SEMI KAZAKH" ~ "Semipalatinsk, Kazakhstan (USSR)",
      region == "STAVRO RUSS" ~ "Stavropol, Russia",
      region == "TUYMEN RUSS" ~ "Tyumen, Russia",
      region == "TYUMEN RUSS" ~ "Tyumen, Russia",
      region == "UKEAINE" ~ "Ukraine (USSR)",
      region == "UKRAINE" ~ "Ukraine (USSR)",
      region == "URAL RUSS" ~ "Ural Mountains, Russia",
      region == "UZBEK" ~ "Uzbekistan, USSR",
      region == "W KAZAKH" ~ "Western Kazakhstan, USSR",
      region == "W MURUROA" ~ "Western Mururoa Atoll, French Polynesia",
      region == "WSW MURUROA" ~ "West-Southwest Mururoa Atoll, French Polynesia",
      TRUE ~ "Unknown/Unmapped Region"  # Default for regions not explicitly listed
    )
  )

# Count the number of bombs per region
top_regions <- mutated_regions %>%
  group_by(region_desc, country) %>%
  summarize(total_bombs = n(), .groups = "drop") %>%  # Count bombs in each region
  arrange(desc(total_bombs)) %>%  # Sort by the number of bombs
  slice_head(n = 10)  # Select the top 10 regions

# Create the bar plot
graph_5 <- top_regions %>%
  ggplot(aes(
    x = reorder(region_desc, total_bombs),  # Order regions by bomb count
    y = total_bombs,
    fill = country,
    text = paste0(
      "Region: ", region_desc,
      "\nCountry: ", country,
      "\nNumber of Bombs: ", total_bombs
    )
  )) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = c(
      "United States" = "#1f77b4",
      "Soviet Union" = "#d62728",
      "United Kingdom" = "#9467bd",
      "France" = "#2ca02c",
      "China" = "#ff7f0e",
      "India" = "#bcbd22",
      "Pakistan" = "#17becf"
    )
  ) +
  labs(
    title = "Top 10 Most Common Regions for Bomb Drops",
    x = "Region",
    y = "Number of Bombs",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16)
  )

interactive_plot <- ggplotly(graph_5, tooltip = "text")
interactive_plot

```

```{r, fig.width=12, fig.height=4}

ggarrange(graph_1, graph_2)

```

