---
title: "code"
author: "Susan Gogolski"
date: "2024-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
Susan’s Code for project

```{r}
library(ggplot2)
library(tidyverse)
library(conflicted)
library(ggridges)
library(plotly)
library(dplyr)
library(RColorBrewer) 
library(forcats)
library(scales)
library(stringr)
```
# Read in nuclear explosions data
```{r}
explosions <- read.csv("./data/nuclear_explosions.csv")
```

# Calculate total number of instances for each purpose
```{r}
purpose <- explosions %>%
  group_by(purpose) %>%
  summarise(number_instances = n(), .groups = "drop")
```
# Select the top 5 purposes bases on the number of instances
```{r}
top_5_purposes <- purpose %>%
  arrange(desc(number_instances)) %>%
  slice_head(n = 5)
```
# Join with the original dataset to get corresponding `country` and `yield_upper`
# Here, we use `left_join` to keep all columns from the original `purpose` data

```{r}
# Join the top 5 purposes with the explosions data
top_5_purposes_data <- explosions %>%
  dplyr::filter(purpose %in% top_5_purposes$purpose) %>%
  # Join with the `top_5_purposes` to get `number_instances` from the purpose dataset
  left_join(top_5_purposes, by = "purpose") %>%
  select(purpose, number_instances, country, yield_upper, year)
```
```{r}
# Replace abbreviations with full words
top_5_purposes_data <-  top_5_purposes_data %>%
  mutate(
    purpose = recode(purpose,
                     "FMS" = "Soviet test",
                     "PNE" = "Peaceful nuclear explosion",
                     "SE" = "Accidental safety testing",
                     "WE" = "Target nuclear detonation",
                     "WR" = "Weapons development")
  )
```

```{r, fig.width=10, fig.height=5}
ggplot(top_5_purposes_data, aes(x = number_instances, y = purpose, fill = purpose, alpha = 0.5)) +
  geom_density_ridges() +
  scale_x_log10() +  # Apply log10 transformation to the x-axis
  scale_fill_manual(values = c("yellow", "green", "purple", "orange", "red")) +  # Custom colors
  scale_x_continuous(labels = label_number()) +
  labs(title = 'Purpose of Nuclear Testing versus Number of Times', x = 'Number of Times', y = 'Purpose of Nuclear Testing') +
  theme_ridges() +
  theme(legend.position = "none",
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5, angle = 90))
```

```





